Rax: Deep Reinforcement Learning for Congestion Control

### 2018-06-21 tex/paper.tex
Final version

\documentclass[newfonts=false,format=sigconf,10pt,letterpaper]{acmart}

@@

% Copyright
\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference info
\setcopyright{none}

@@

\begin{document}
\title{TCP Rax: Deep Reinforcement Learning for Congestion Control}
%\titlenote{Produces the permission block, and
%  copyright information}
%\subtitle{Extended Abs\texttt{tract}
%\subtitlenote{The full version of the author's guide is available as
%  \texttt{acmart.pdf} document}}

@@

A major challenge in TCP congestion control is to find a set of rules that dynamically adapt the congestion window to current and anticipated future network conditions to maximize a specified utility function, e.g. maximum throughput or minimum packet loss.

This paper proposes Reactive Adaptive eXperience based congestion control (TCP Rax), a new method of congestion control that uses online reinforcement learning to maintain an optimum congestion window with respect to a given utility function and based on current network conditions. We use a neural network based approach that can be initialized either with random weights (without prior knowledge) or with a previously trained neural network to improve stability and convergence time. The processing of rewards in congestion control depends on the arrival of acknowledgments, inducing loop delays that may lead to oscillations. As a remedy we propose \textit{Partial Action Learning} (PAL), a specific formulation of neural network based reinforcement learning that supports the processing of delayed and partial rewards.

We show that our method converges to a stable, close-to-optimum solution within the order of minutes and outperforms existing congestion control algorithms in typical networks. Thus, this paper demonstrates that neural network based reinforcement learning without any prior knowledge can feasibly be done online and can compete with human-designed congestion-control schemes such as New Reno and Cubic.

@@

Recent advances in neural network based reinforcement learning (RL) that for the first time have the ability to learn complex tasks such as video games, raise the question if it is possible to use these frameworks to model end-to-end congestion control: As in RL, in congestion control each sender has a an observable state, performs actions and receives rewards for its actions according to the sender's objective function.

@@

To address these problems, we develop a new formulation of reinforcement learning called \textit{Partial Action Learning} (PAL). PAL is a superset of reinforcement learning: If one uses PAL for a learning problem without delay, asynchronicity and partial actions/partial rewards, one gets classical reinforcement learning.

@@

\begin{figure}
\begin{minipage}{\columnwidth}
\subfloat[One sender\label{fig:demo1}]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/4_1_2018-01-25-20:17:15_3_2018-01-25-20:17:15.0}}}}{}
\subfloat[Two senders\label{fig:demo2}]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-6.local/figures/4_2_2018-01-25-19:53:39_22_2018-01-25-19:53:39.0}}}}{}
\caption{Throughput and loss rate when sharing a bottleneck link of 20\,Mbit/s with a two-way end-to-end delay of 50\,ms and a very small buffer of $\frac{1}{10}$ bandwidth delay product between one (\ref{fig:demo1}) and two (\ref{fig:demo2}) senders . The time needed for convergence increases with more senders, which is caused by the fact that the more senders use the link, the fewer packets each sender receives per unit of time for training and because of the increasing complexity of the environment. In case of two senders, both flows complement each other and quickly adapt to the sending rate of the other flow.}
\label{fig:demo}
\end{minipage}
\end{figure}

@@

Congestion control has been implemented in TCP since the 1980s as a series of congestion collapses dramatically decreased applications' throughput in the Internet, which lead to the development of the Tahoe and Reno algorithms \cite{jacobson_congestion_1988}. These approaches as well as most others maintain a congestion window which stands for the maximum amount of data that are allowed to be in flight in the network without having been acknowledged by the receiver. In most TCP congestion control algorithms, receiving acknowledgements of data that were successfully transmitted leads to the congestion window being increased while packet loss (or an increase in delay as in the Vegas algorithm \cite{brakmo_tcp_1995}) leads to the congestion window being lowered according to a set of fixed rules. Algorithms such as Cubic and Compound \cite{ha_cubic:_2008, tan_compound_2006} improve congestion control specifically with respect to so-called long fat networks with a high round-trip time and bandwidth. Proportional rate reduction \cite{dukkipati_proportional_2011} aims to make the reduction of the rate that occurs in case of congestion more smooth and steady in time to avoid the bursty behavior that occurred upon loss in previous TCP congestion control variants. A different approach is taken by BBR \cite{cardwell_bbr:_2016}, which does not use a set of fixed rules like previous congestion control algorithms. Instead it estimates a model of the network path by using measurements of RTT and throughput and then adjusts its sending rate so that it uses the maximum bandwidth according to its network model while trying not to fill up queues. A similar approach is PCC \cite{dong_pcc:_2015}, which also uses measurements to find an optimum sending rate with respect to a defined utility function (e.g. maximizing throughput while minimizing packet loss). However, PCC doesn't maintain a congestion window but instead modulates the sending rate, using one constant sending rate for a certain time span.

@@

To this end we propose Partial Action Learning which is based on the Asynchronous Advantage Actor Critic framework \cite{mnih_asynchronous_2016}, which has been demonstrated to be able to learn to play a wide range of video games and commonly outperform human players. Its use of a neural network enables it to even learn complex tasks. In particular, it is a good choice for congestion control as it has been proven to deliver good performance and high training speed for playing video games and also for networking related tasks such as Adaptive Bitrate Algorithms \cite{mao_neural_2017}.

@@

Our proposed framework \textit{Partial Action Learning} (PAL) (see \ref{subsec:pal}) is based on the Actor Critic framework for neural networks proposed by \citet{mnih_asynchronous_2016}, which we outline (with a focus on intuition over rigor) in this section.

@@

To compute $R_t$ at time step $t$ it is necessary to look infinitely far into the future to get all future rewards as in each step $t$ computing $R_{t+1}$ relies on knowing $R_{t+1}$. Thus, the algorithm takes $t_\text{max}$ steps (for example 20), waits for all the rewards and then calculates $R_t$ for each step and uses an estimate of $R_{t_\text{max}+1}$ provided by the critic network as the continuation in the future. The only exception from this is when the task to execute ends before reaching $t_\text{max}$ (e.g. the TCP flow is over). In this case, one can either use a prediction from the critic network as a continuation or alternatively let the moving average end without continuation i.e.~ignore the term weighted with $(\gamma-1)$ at the last step.

@@

The key difference between our approach PAL and previous approaches to RL is that in classical RL an action is always followed by a reward and a reward is always followed by an action. In our proposed concept, however, it is possible to take new actions while previous actions haven't been rewarded yet.

Another major difference in PAL is that one action generates a number of partial actions ($\geq 0$) (see \autoref{fig:pal}). Each partial action generates feedback upon interacting with the environment. Upon receiving feedback for a partial action, the agent determines the current state and performs a new action. When all feedbacks of one action were received, the agent combines them to form the reward and updates the critic and actor networks.

In \autoref{alg:pal} (in the appendix) we show the code that runs in each of the agents (in the congestion control scenario, one agent corresponds to a sender). It is possible to have several agents which share a set of neural weights (which is usually done when performing offline training \cite{mnih_asynchronous_2016,mao_neural_2017}) but one can also use separate weights for each agent, which is more realistic in case of congestion control in the Internet, as it is not sensible that different senders share a set of neural network weights over the Internet.

@@

We use a reward function that is similar to the one used by PCC \cite{dong_pcc:_2015}.

@@

The tunable tradeoff between bandwidth and packet loss rate ($\alpha$) has a clear impact on both the highest spike of packet loss that is produced in the very beginning of the online learning process as well as on the median packet loss rate later on in the learning process (see \autoref{fig:one_sender_loss}). However, the stronger packet loss is punished, the longer it takes for the sender to reach the maximum bandwidth on the link.

@@

While TCP Rax generally does not exhibit much variance after a few minutes of training when it is the only flow on a link, fluctuations and training time needed to achieve good performance increase when having two or four concurrent senders (see \autoref{fig:multiple_senders}). We attribute this to the fact that on a shared link with $n$ senders each can only get $\frac{1}{n}$ of the bandwidth on average and thus each sender receives fewer packets which it can use for training and therefore training takes longer. Furthermore, as the environment becomes more dynamic with more senders, each sender has a harder time learning whether the reward changed due to his own actions or due to actions that were caused by other senders: If there is only one flow on a link, the flow ``can be sure'' that all changes in rewards were actually caused by his own actions.

@@

To verify whether pre-training is generally feasible, we used the neural networks that were produced after each 600 second flow shown in \autoref{fig:multiple_senders} and started flows of equal length with the pre-trained neural networks. The flows seem to simply resume where they ended training in the flow before (see \autoref{fig:resuming_training}). This does not seem like a remarkable finding, however, it is quite unexpected: For instance, in the case of one sender on a link, the neural network starts with random weights and discovers that it achieves optimum utility at a window that results in 20\,Mbit/s of bandwidth. Then the sender constantly stays in this state and one would expect it to slowly ``forget'' what it learned in the very beginning when it started the flow. Thus, it is interesting that after hundreds of seconds of only seeing one same state, the neural network still acts correctly when starting with a small congestion window again in a completely different state than the one it was used to for hundreds of seconds before. For two senders pre-training also greatly speeds up the initial training procedure and can avoid the spike of packet loss in the beginning that always occurs to some extent without pre-training (see \autoref{fig:one_sender_loss} and \autoref{fig:multiple_senders}). Pre-training also works for more than one flow and although TCP Rax learns to minimize packet loss even further (see Figure \ref{fig:multiple_senders2}) some variability in throughput remains. We argue that this is caused by the stochastic behavior of the reinforcement learning: While each sender aims to maximize its utility, at the same time it also tries out new actions to check whether they could improve its utility even further and this behavior results in fluctuations.

@@

For reasons of comparability we take the same scenario of a 20\,Mbit/s shared bottleneck with a small buffer of $\frac{1}{10}$ bandwidth delay product to create a challenging scenario. Surprisingly, New Reno performs by far better than more modern TCP congestion control algorithms (see \autoref{fig:regular_tcps}). We attribute this to the fact that we chose a small buffer to create a challenging scenario and that Cubic and Compound are not optimized so well for this scenario but rather for long fat networks \cite{oura_fairness_2012}. TCP Rax achieves a significantly larger total throughput on average though with a couple of outliers which become more significant with two concurrent senders (see \autoref{fig:rax_tcps}); when inspecting the raw data, it becomes apparent that these outliers occur usually when the neural network diverges in the first seconds of a flow.

@@

Our results demonstrate that our proposed concept can perform online learning of congestion control and can outperform existing congestion control algorithms in certain scenarios. Our results also show that with an increasing number of senders the reinforcement learning becomes significantly more challenging and time consuming. Regarding online learning, during the implementation of TCP Rax it became apparent that a neural network consisting of LSTM cells \cite{gers_learning_1999} is not feasible. LSTM cells keep an internal state that gets unstable when it is used for long training episodes, as for each received acknowledgement the LSTM state has to be updated. Previous work using deep reinforcement learning uses up to 50 million training steps for offline learning \cite{mnih_asynchronous_2016} while we generally use significantly less (a 20 Mbit/s link can carry a maximum of up to one million packets of 1250 bytes size in 10 minutes). When sharing the link between several senders, each sender receives only a fraction of these packets for training. For example, with two senders on one link (see Figure \ref{fig:multiple_senders2}) one sender can receive a maximum of 600\,000 packets during a ten minute flow for training. We argue that the reason for TCP Rax needing fewer training samples than other reinforcement learning tasks is that in congestion control rewards are steadily coming in, while for other tasks, such as video games, commonly rewards are only received sparsely (e.g. one has to wait until the game is over to determine whether one won or lost the game).

@@

\begin{algorithm}
\caption{Partial Action Learning -- procedure which computes and applies the gradients.}
\label{alg:grad}
\begin{algorithmic}[1]%
%\algrestore{palalg}
\Function{computeGradients}{}
	\State $t_\text{end} = \#\left( l_\text{rewards} \right)$
	\State $\theta_\text{backup} \gets \theta$
	\State $h_\text{backup} \gets h$
	\State $\theta \gets l_\text{snapshots}[0]$
	\State $h \gets l_\text{hiddenStates}[0]$
	\State $R_{i+1} \gets$ last element of $l_\text{values'}$%
	\For{$i\gets t_\text{end}-1..0$}
		\State $R_i \gets \left(r_i + \gamma R_{i+1}\right)\left(1-\gamma\right)$
		\State $a \gets l_\text{actions}[i]$
		\State $s \gets l_\text{states}[i]$
		\State $v \gets l_\text{values}[i]$
		\State $d\theta \gets -\frac{\partial\log \left( \pi\left( a \given s; \theta_\text{a} \right)\right)\left(R_i - v\right)}{\partial\theta}$
		\StateIndent[4] $-\frac{\partial\beta H\left( \pi\left( s; \theta_a \right)\right)}{\partial\theta}$
		\StateIndent[4] $+\frac{\partial\left(R_i - v\right)}{\partial\theta}$
		\State $\theta_{\text{g}} \gets \theta_{\text{g}} + d\theta$
	\EndFor
	\State Remove first $t_\text{end}$ elements from
	\StateIndent[2] $l_\text{actions}$, $l_\text{states}$, $l_\text{values}$, $l_\text{rewards}$, $l_\text{values'}$
	\State Remove the first element from $l_\text{snapshots}$, $l_\text{hiddenStates}$
	\State $\theta = \theta_\text{backup}$
	\State $h = h_\text{backup}$
\EndFunction
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{Procedure which determines when to start computing the gradients in case of congestion control. We assume that the window at time steps $t$ is stored in a list $l_\text{windows}$ and that the number of bytes sent in each time step is stored in $l_\text{sent}$ analogously to the other metrics defined in \autoref{alg:pal}. Alternatively to this procedure, one can simply compute the gradients when $t_\text{max}$ rewards have been received where $t_\text{max}$ is an arbitrarily chosen integer greater than 0.}
\label{alg:enough}
\begin{algorithmic}[1]%
%\algrestore{palalg}
\Function{gotEnoughRewards?}{}
	\If{the epsiode is over}
		\State \Return \True
	\EndIf
	\State $w \gets \left\lfloor l_\text{windows}[0] + l_\text{actions}[0] \right\rfloor$
	\For{$i\gets 0..\#(l_\text{rewards})-1$}
		\State $w \gets \text{w} - l_\text{sent}[i]$
		\If{$w \leq 0 \land i+1 \geq 10$} \Comment{at least 10}
			\State \Return \True
		\EndIf
	\EndFor
	\State \Return \False
\EndFunction
\end{algorithmic}
\end{algorithm}

>>>>

\documentclass[10pt,sigconf,anonymous]{acmart}

@@

%% Copyright
%\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference info
%\setcopyright{none}

%Conference Info
\acmYear{2018}
\copyrightyear{2018}
\setcopyright{acmcopyright}
\acmConference{CoNEXT '18}{December 4--7, 2018}{Heraklion/Crete, Greece}
\acmPrice{TBA}
\acmDOI{TBA}
\acmISBN{TBA}

@@

\begin{document}
\title{Rax: Deep Reinforcement Learning for Congestion Control}
%\titlenote{Produces the permission block, and
%  copyright information}
%\subtitle{Extended Abs\texttt{tract}
%\subtitlenote{The full version of the author's guide is available as
%  \texttt{acmart.pdf} document}}

@@

A major challenge in congestion control is to find a set of rules that dynamically adapt the congestion window to current and anticipated future network conditions to maximize a specified utility function, e.g. maximum throughput or minimum packet loss.

This paper proposes Reactive Adaptive eXperience based congestion control (Rax), a new method of congestion control that uses online reinforcement learning to maintain an optimum congestion window with respect to a given utility function and based on current network conditions. We use a neural network based approach that can be initialized either with random weights (without prior knowledge) or with a previously trained neural network to improve stability and convergence time. As the processing of rewards in congestion control depends on the arrival of acknowledgments, which are delayed and come piece by piece, the problem is not suitable for current implementations of deep reinforcement learning. As a remedy we propose \textit{Partial Action Learning} (PAL), a specific formulation of deep reinforcement learning that supports the processing of delayed and partial rewards.

We show that our method converges to a stable, close-to-optimum solution within the order of minutes and outperforms existing congestion control algorithms in typical networks. Thus, this paper demonstrates that neural network based reinforcement learning can be done online and can compete with human-designed congestion-control schemes such as New Reno and Cubic.

@@

\input{concept.tex}

\caption{Hybrid learning approach to congestion control: After initial training in a simulator (Offline Learning), the congestion controller is deployed in a real network where it performs congestion control while at the same time still learning ways to improve (Online Learning). After sufficient training it is also possible to disable the online learning and use the previously learned knowledge (Static Operation). This can help reduce computational resources.}
\label{fig:concept}
\end{figure}

Congestion control has been a long-standing problem in networking. In the past decades various attempts have been made to devise a set of rules that performs optimum congestion control. More recently, researchers have realized that such a set of rules might be impossible to find due to the complexity of the current Internet (caused by various types middleboxes, different congestion control algorithms and Active Queue Management) and the fast-paced evolution of Internet technologies. Two principal solutions have been proposed recently: \begin{enumerate*}
\item Remy, a machine learning approach that is trained offline in a simulator and then statically deployed \cite{winstein_tcp_2013}, and \item PCC, which learns online but forgets all it learned after each flow \cite{dong_pcc:_2015} \end{enumerate*}. We argue that the solution lies in a hybrid approach: A machine learning based congestion control that receives some initial pre-training and then uses online learning to constantly refine and optimize its behavior (\autoref{fig:concept}).

Recent advances in neural network based reinforcement learning (RL), which for the first time have the ability to learn complex tasks such as video games, seem to be an ideal tool to model end-to-end congestion control: As in RL, in congestion control each sender has a an observable state, performs actions and receives rewards for its actions according to the sender's objective function.

@@

To address these problems, we develop a new implementation of deep reinforcement learning called \textit{Partial Action Learning} (PAL) which we base on the Asynchronous Advantage Actor Critic framework \cite{mnih_asynchronous_2016}.

@@

\begin{figure}
\begin{minipage}{\columnwidth}
\subfloat[One sender\label{fig:demo1}]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/4_1_2018-01-25-20:17:15_3_2018-01-25-20:17:15.0}}}}{}
\subfloat[Two senders\label{fig:demo2}]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-6.local/figures/4_2_2018-01-25-19:53:39_22_2018-01-25-19:53:39.0}}}}{}
\caption{Throughput and loss rate when sharing a bottleneck link of 20\,Mbit/s with a two-way end-to-end delay of 50\,ms and a very small buffer of $\frac{1}{10}$ bandwidth delay product between one (\ref{fig:demo1}) and two (\ref{fig:demo2}) senders . The time needed for convergence increases with more senders, because the more senders use the link, the fewer packets each sender receives per unit of time for training and because of the increasing complexity of the environment. In case of two senders, both flows complement each other and quickly adapt to the sending rate of the other flow.}
\label{fig:demo}
\end{minipage}
\end{figure}

@@

Early congestion control algorithms such as Tahoe and Reno \cite{jacobson_congestion_1988} pioneered the idea to maintain a congestion window which stands for the maximum amount of data that are allowed to be in flight in the network without having been acknowledged by the receiver. In most TCP congestion control algorithms, receiving acknowledgements of data that were successfully transmitted leads to the congestion window being increased while packet loss (or an increase in delay as in the Vegas algorithm \cite{brakmo_tcp_1995}) leads to the congestion window being lowered according to a set of fixed rules. Algorithms such as Cubic and Compound \cite{ha_cubic:_2008, tan_compound_2006} improve congestion control specifically with respect to so-called long fat networks with a high round-trip time and bandwidth. A different approach is taken by BBR \cite{cardwell_bbr:_2016}, which does not use a set of fixed rules like previous congestion control algorithms. Instead it estimates a model of the network path by using measurements of RTT and throughput and then adjusts its sending rate so that it uses the maximum bandwidth according to its network model while trying not to fill up queues. A similar approach is PCC \cite{dong_pcc:_2015}, which also uses measurements to find an optimum sending rate with respect to a defined utility function (e.g. maximizing throughput while minimizing packet loss). However, PCC doesn't maintain a congestion window but instead modulates the sending rate, using one constant sending rate for a certain time span.

@@

To this end we propose Partial Action Learning which is based on the Asynchronous Advantage Actor Critic framework \cite{mnih_asynchronous_2016}, which has been demonstrated to be able to learn to play a wide range of video games and commonly outperform human players. Its use of a neural network enables it to even learn complex tasks. In particular, it is a good choice for congestion control as it has been proven to deliver good performance and high training speed for not only playing video games but also for networking related tasks such as Adaptive Bitrate Algorithms \cite{mao_neural_2017} and Traffic Engineering \cite{xu_experience-driven_2018}.

@@

Actor Critic methods of reinforcement learning were first studied by \cite{sutton_policy_1999} and recently became popular for deep learning. Our proposed framework \textit{Partial Action Learning} (PAL) (see \ref{subsec:pal}) is based on the Actor Critic framework for neural networks proposed by \citet{mnih_asynchronous_2016}, which we outline in this section.

@@

To compute $R_t$ at time step $t$ it is necessary to look infinitely far into the future to get all future rewards as in each step $t$ computing $R_{t+1}$ relies on knowing $R_{t+1}$. Thus, the algorithm takes $t_\text{max}$ steps (for example 20), waits for all the rewards and then calculates $R_t$ for each step and uses an estimate of $R_{t_\text{max}+1}$ provided by the critic network as the continuation in the future. The only exception from this is when the task to execute ends before reaching $t_\text{max}$ (e.g. the TCP flow is over). In this case, one can let the moving average end without continuation i.e.~ignore the term weighted with $(\gamma-1)$ at the last step.

@@

The key difference between our approach PAL and previous implementations of deep RL is that in our approach we can handle the delay of rewards. Another major difference in PAL is that one action generates a number of partial actions ($\geq 0$) (see \autoref{fig:pal}). Each partial action generates feedback upon interacting with the environment. Upon receiving feedback for a partial action, the agent determines the current state and performs a new action. When all feedback of one action was received, the agent combines it to form the reward and updates the critic and actor networks.

In \autoref{alg:pal} (in the appendix) we show the code that runs in each of the agents (in the congestion control scenario, one agent corresponds to a sender). It is possible to have several agents which share a set of neural network weights (which is usually done when performing offline training \cite{mnih_asynchronous_2016,mao_neural_2017}) but one can also use separate weights for each agent, which is more realistic in case of congestion control in the Internet, as it is not sensible that different senders share a set of neural network weights over the Internet.

@@

We use a reward function that is similar to the one used by PCC, whose convergence for multiple senders has been proven previously \cite{dong_pcc:_2015}.

@@

The tunable tradeoff between bandwidth and packet loss rate ($\alpha$) has a clear impact on both the highest spike of packet loss that is produced in the very beginning of the online learning process as well as on the median packet loss rate later on in the learning process (\autoref{fig:one_sender_loss}). However, the stronger packet loss is punished, the longer it takes for the sender to reach the maximum bandwidth on the link.

@@

While TCP Rax generally does not exhibit much variance after a few minutes of training when it is the only flow on a link, fluctuations and training time needed to achieve good performance increase when having two or four concurrent senders (\autoref{fig:multiple_senders}). We attribute this to the fact that on a shared link with $n$ senders each can only get $\frac{1}{n}$ of the bandwidth on average and thus each sender receives fewer packets which it can use for training and therefore training takes longer. Furthermore, as the environment becomes more dynamic with more senders, each sender has a harder time learning whether the reward changed due to his own actions or due to actions that were caused by other senders: If there is only one flow on a link, the flow ``can be sure'' that all changes in rewards were actually caused by his own actions.

@@

To verify whether pre-training is generally feasible, we used the neural networks that were produced after each 600 second flow shown in \autoref{fig:multiple_senders} and started flows of equal length with the pre-trained neural networks. The flows seem to simply resume where they ended training in the flow before (\autoref{fig:resuming_training}). This does not seem like a remarkable finding, however, it is quite unexpected: For instance, in the case of one sender on a link, the neural network starts with random weights and discovers that it achieves optimum utility at a window that results in 20\,Mbit/s of bandwidth. Then the sender constantly stays in this state and one would expect it to slowly ``forget'' what it learned in the very beginning when it started the flow. Thus, it is interesting that after hundreds of seconds of only seeing one same state, the neural network still acts correctly when starting with a small congestion window again in a completely different state than the one it was used to for hundreds of seconds before. For two senders pre-training also greatly speeds up the initial training procedure and can avoid the spike of packet loss in the beginning that always occurs to some extent without pre-training (\autoref{fig:one_sender_loss} and \autoref{fig:multiple_senders}). Pre-training also works for more than one flow and although TCP Rax learns to minimize packet loss even further (Figure \ref{fig:multiple_senders2}) some variability in throughput remains. We argue that this is caused by the stochastic behavior of the reinforcement learning: While each sender aims to maximize its utility, at the same time it also tries out new actions to check whether they could improve its utility even further and this behavior results in fluctuations.

@@

For reasons of comparability we take the same scenario of a 20\,Mbit/s shared bottleneck with a small buffer of $\frac{1}{10}$ bandwidth delay product to create a challenging scenario. Surprisingly, New Reno performs by far better than more modern TCP congestion control algorithms (\autoref{fig:regular_tcps}). We attribute this to the fact that we chose a small buffer to create a challenging scenario and that Cubic and Compound are not optimized so well for this scenario but rather for long fat networks \cite{oura_fairness_2012}. TCP Rax achieves a significantly larger total throughput on average though with a couple of outliers which become more significant with two concurrent senders (\autoref{fig:rax_tcps}); when inspecting the raw data, it becomes apparent that these outliers occur usually when the neural network diverges in the first seconds of a flow.

@@

Our results demonstrate that our proposed concept can perform online learning of congestion control and can outperform existing congestion control algorithms in certain scenarios. Our results also show that with an increasing number of senders the reinforcement learning becomes significantly more challenging and time consuming. Regarding online learning, during the implementation of TCP Rax it became apparent that a neural network consisting of LSTM cells \cite{gers_learning_1999} is not feasible. LSTM cells keep an internal state that gets unstable when it is used for long training episodes, as for each received acknowledgement the LSTM state has to be updated. Previous work using deep reinforcement learning uses up to 50 million training steps for offline learning \cite{mnih_asynchronous_2016} while we generally use significantly less (a 20 Mbit/s link can carry a maximum of up to one million packets of 1250 bytes size in 10 minutes). When sharing the link between several senders, each sender receives only a fraction of these packets for training. For example, with two senders on one link (Figure \ref{fig:multiple_senders2}) one sender can receive a maximum of 600\,000 packets during a ten minute flow for training. We argue that the reason for TCP Rax needing fewer training samples than other reinforcement learning tasks is that in congestion control rewards are steadily coming in, while for other tasks, such as video games, commonly rewards are only received sparsely (e.g. one has to wait until the game is over to determine whether one won or lost the game).

@@

%\begin{algorithm}
%\caption{Partial Action Learning -- procedure which computes and applies the gradients.}
%\label{alg:grad}
%\begin{algorithmic}[1]%
%%\algrestore{palalg}
%\Function{computeGradients}{}
%	\State $t_\text{end} = \#\left( l_\text{rewards} \right)$
%	\State $\theta_\text{backup} \gets \theta$
%	\State $h_\text{backup} \gets h$
%	\State $\theta \gets l_\text{snapshots}[0]$
%	\State $h \gets l_\text{hiddenStates}[0]$
%	\State $R_{i+1} \gets$ last element of $l_\text{values'}$%
%	\For{$i\gets t_\text{end}-1..0$}
%		\State $R_i \gets \left(r_i + \gamma R_{i+1}\right)\left(1-\gamma\right)$
%		\State $a \gets l_\text{actions}[i]$
%		\State $s \gets l_\text{states}[i]$
%		\State $v \gets l_\text{values}[i]$
%		\State $d\theta \gets -\frac{\partial\log \left( \pi\left( a \given s; \theta_\text{a} \right)\right)\left(R_i - v\right)}{\partial\theta}$
%		\StateIndent[4] $-\frac{\partial\beta H\left( \pi\left( s; \theta_a \right)\right)}{\partial\theta}$
%		\StateIndent[4] $+\frac{\partial\left(R_i - v\right)}{\partial\theta}$
%		\State $\theta_{\text{g}} \gets \theta_{\text{g}} + d\theta$
%	\EndFor
%	\State Remove first $t_\text{end}$ elements from
%	\StateIndent[2] $l_\text{actions}$, $l_\text{states}$, $l_\text{values}$, $l_\text{rewards}$, $l_\text{values'}$
%	\State Remove the first element from $l_\text{snapshots}$, $l_\text{hiddenStates}$
%	\State $\theta = \theta_\text{backup}$
%	\State $h = h_\text{backup}$
%\EndFunction
%\end{algorithmic}
%\end{algorithm}

@@

%\begin{algorithm}
%\caption{Procedure which determines when to start computing the gradients in case of congestion control. We assume that the window at time steps $t$ is stored in a list $l_\text{windows}$ and that the number of bytes sent in each time step is stored in $l_\text{sent}$ analogously to the other metrics defined in \autoref{alg:pal}. Alternatively to this procedure, one can simply compute the gradients when $t_\text{max}$ rewards have been received where $t_\text{max}$ is an arbitrarily chosen integer greater than 0.}
%\label{alg:enough}
%\begin{algorithmic}[1]%
%%\algrestore{palalg}
%\Function{gotEnoughRewards?}{}
%	\If{the epsiode is over}
%		\State \Return \True
%	\EndIf
%	\State $w \gets \left\lfloor l_\text{windows}[0] + l_\text{actions}[0] \right\rfloor$
%	\For{$i\gets 0..\#(l_\text{rewards})-1$}
%		\State $w \gets \text{w} - l_\text{sent}[i]$
%		\If{$w \leq 0 \land i+1 \geq 10$} \Comment{at least 10}
%			\State \Return \True
%		\EndIf
%	\EndFor
%	\State \Return \False
%\EndFunction
%\end{algorithmic}
%\end{algorithm}
